---
title: "CS 422 HW5"
author: "Ana Velasco"
output:
  html_notebook:
    toc: yes
    toc_float: yes
  html_document:
    toc: yes
    df_print: paged
---

### Libraries
```{r}
library(arules)
library(arulesViz)
library(lsa)
library(dplyr, verbose=F)
rm(list=ls())
```

### Part 2.1-A.

```{r}
setwd("C:/Users/anave/OneDrive - Universidad Politécnica de Madrid/MUIT/ERASMUS/1_PRIMER_CUATRI/CS422DATAMINING/Assignment5")

products <- read.csv("products.csv", header=F)

for (i in c(1, 5, 20, 75)){
  name <- paste("tr-",i,"k.csv", sep = "")
  products.df <-read.csv(name, header=F, fill=T)[,-1]
  
  can<- function(x){
    x <- products$V2[match(x,products$V1)]
  }
  products.can <-data.frame(apply(products.df,2, FUN=can))
  colnames(products.can) <- NULL
  
  name.can <-paste("tr-", i, "k-canonical.csv", sep="")
  write.table(products.can, name.can, row.names=FALSE, quote=FALSE, sep=",", na="")
}
```
### Part 2.1-B
In this section I will import the transactions from the file created in section 2.1-A. 

Then, I will find the frequent itemsets. An itemset is frequent if the support of that itemset is greater or equal than the minsup (minimum support). For all the 4 data sets, this value has been achieved by gradually descending from high values, until sufficient patterns and about 20 strong rules have been found. 

Finally I will discover association rules within the frequent itemsets. For that purpose, rules are inspected sorting them by confidence. 
```{r}
# 1K Transactions
#Frequent itemsets
trans1k <- read.transactions("tr-1k-canonical.csv", sep=",")
summary(trans1k)
freq.itemsets1k <- apriori(trans1k, parameter=list(support=0.027, target="frequent itemsets"), control=list(verbose=F))
inspect(sort(freq.itemsets1k, by="count", decreasing=T))

#Association rules
rules1k <- apriori(trans1k, parameter=list(support=0.027, conf=0.75), control=list(verbose=F))
inspect(sort(rules1k, by="confidence", decreasing=T))

# 5K Transactions
#Frequent itemsets
trans5k <- read.transactions("tr-5k-canonical.csv", sep=",")
freq.itemsets5k <- apriori(trans5k, parameter=list(support=0.024, target="frequent itemsets"), control=list(verbose=F))
inspect(sort(freq.itemsets5k, by="count", decreasing=T))

#Association rules
rules5k <- apriori(trans5k, parameter=list(support=0.024, conf=0.75), control=list(verbose=F))
inspect(sort(rules5k, by="confidence", decreasing=T))

# 20K Transactions
#Frequent itemsets
trans20k <- read.transactions("tr-20k-canonical.csv", sep=",")
freq.itemsets20k <- apriori(trans20k, parameter=list(support=0.027, target="frequent itemsets"), control=list(verbose=F))
inspect(sort(freq.itemsets20k, by="count", decreasing=T))
#Association rules
rules20k <- apriori(trans20k, parameter=list(support=0.027, conf=0.75), control=list(verbose=F))
inspect(sort(rules20k, by="confidence", decreasing=T))

# 75K Transactions
#Frequent itemsets
trans75k <- read.transactions("tr-75k-canonical.csv", sep=",")
freq.itemsets75k <- apriori(trans75k, parameter=list(support=0.019, target="frequent itemsets"), control=list(verbose=F))
inspect(sort(freq.itemsets75k, by="count", decreasing=T))
#Association rules
rules75k <- apriori(trans75k, parameter=list(support=0.019, conf=0.75), control=list(verbose=F))
inspect(sort(rules75k, by="confidence", decreasing=T))

```

### Part 2.1-C

It can clearly be seen that the value of minsup variates depending on the data set. 
Both data sets with 1K and 20K transactions needed a higher minsup (0.027) in order to obtain a confidence higher than 0.75 (about 20 strong rules).
However, both 5k and 75k data sets, required a lower minsup (0.024 and 0.019 respectively)

### Part 2.1-D-i

The most frequently purchased item/itemset in the 75K transactions data set, it is purchased in 8191 transactions and corresponds to {Coffee Eclair}. 

### Part 2.1-D-ii

The least frequently purchased item/itemset in the 75K transactions data set, it is purchased in 1744 transactions and corresponds to {Apple Croissant, Apple Danish,Apple Tart}
