---
title: "Homework 7"
author: "Ana Velasco"
output:
   html_notebook:
      toc: yes
      toc_float: yes
   html_document:
      toc: yes
      df_print: paged
---
```{r}
library(keras)
library(dplyr)
library(caret)
library(rpart)
library(rpart.plot)

rm(list=ls())
```

# Set working directory as needed
```{r}
setwd("C:/Users/anave/OneDrive - Universidad Politécnica de Madrid/MUIT/ERASMUS/1_PRIMER_CUATRI/CS422DATAMINING/Assignment7")

df <- read.csv("wifi_localization.csv")
```
# Seed the PRNG
```{r}
set.seed(1122)
df <- df[sample(nrow(df)), ] # Shuffle, as all of the data in the .csv file
                             # is ordered by label!  
index <- sample(1:nrow(df), 0.20*nrow(df))
test.df  <- df[index, ]
train.df <- df[-index, ]

```

# --- Your code goes below ---
# (a)
```{r}
model <- rpart( room~ .,data=df,method= "class")
rpart.plot(model, extra=104, fallen.leaves=T, type=4, main="WiFi Access Point Decision Tree")

prediction <- predict(model, newdata=test.df, type="class")
conMatrix <- confusionMatrix(prediction, as.factor(test.df$room))
conMatrix

cat("Decision tree model", "\n")
cat("\t", "Sensitivity Class1:", conMatrix$byClass[1,1], "Class2:",conMatrix$byClass[2,1] , "\n","\t","\t", "    Class3:",conMatrix$byClass[3,1], "Class4:", conMatrix$byClass[4,1], "\n" )
cat("\t", "Specificity Class1:", conMatrix$byClass[1,2], "Class2:",conMatrix$byClass[2,2] , "\n","\t","\t", "    Class3:",conMatrix$byClass[3,2], "Class4:", conMatrix$byClass[4,2], "\n" )
cat("\t", "PPV         Class1:", conMatrix$byClass[1,3], "Class2:",conMatrix$byClass[2,3] , "\n","\t","\t", "    Class3:",conMatrix$byClass[3,3], "Class4:", conMatrix$byClass[4,3], "\n" )
cat("\t", "Bal. Acc.   Class1:", conMatrix$byClass[1,11], "Class2:",conMatrix$byClass[2,11] , "\n","\t","\t", "    Class3:",conMatrix$byClass[3,11], "Class4:", conMatrix$byClass[4,11], "\n" )

```
# (b)
# Note that in (b) either use a new variable to store the model, or null out
# the variable that stored the model in (a) if you want to reuse that variable.
# The reason is that if you don't null it out, the model in (b) will have
# residual information left over from (a) and your results will not be quite
# accurate.
```{r}

x_train <- select(train.df, -room)
y_train <- train.df$room
y_train.ohe <- to_categorical(y_train)
x_test <- select(test.df, -room)
y_test <- test.df$room
y_test.ohe <- to_categorical(y_test)

#Creation, train and evaluation
model_nn <- NULL
create_model <- function(){
  model_nn <- keras_model_sequential()
  model_nn %>%
    layer_dense(units = 1, activation = 'relu', input_shape = c(7)) %>%
    layer_dense(units = 5, activation = 'softmax')

model_nn %>% compile(
  optimizer = 'adam', 
  loss = 'categorical_crossentropy',
  metrics = c('accuracy'))

begin <- NULL
end <- NULL
begin <- Sys.time()

acc<- model_nn %>% fit(data.matrix(x_train), y_train.ohe, epochs = 100, batch_size= 32, validation_split=0.20, verbose=0)

end <- Sys.time()
time.to.train <- end-begin

evaluation <- model_nn %>% evaluate(as.matrix(x_test), y_test.ohe)

out<-list(model_nn, time.to.train, acc,evaluation)
return(out)
}
out <- create_model()

model_nn <- out[[1]]
accur <- out[[3]]
eval <- out[[4]]

```
#2.1 b) i)
```{r}
cat("For one neuron in hidden layer, loss:", eval[[1]], "Accuracy:", eval[[2]], "\n")
```
#2.1 b) ii)

Model accuracy is low because there is only one hidden layer and one neuron in the hidden layer. Adding more hidden layers can help improve the accuracy of a model but only up to a certain point, or instead, adding more neurons would help.

```{r}
# Plot Model Accuracy
plot(accur[["metrics"]][["accuracy"]], xlab="Epochs", ylab="Accuracy", main="Model Accuracy", ylim=c(0,1), col="blue", pch=20, type="o")
points(accur[["metrics"]][["val_accuracy"]], col="grey", pch=20, type="o")
legend("topright", legend=c("train", "test"),col=c("blue", "grey"), pch=c(20,20))
#Plot Model Loss
plot(accur[["metrics"]][["loss"]], xlab="Epochs", ylab="Loss", main="Model Loss", col="blue", pch=20, type="o")
points(accur[["metrics"]][["val_loss"]], col="grey", pch=20, type="o")
legend("topright", legend=c("train", "test"),col=c("blue", "grey"), pch=c(20,20))

#Confusion Matrix
# Prediction
pred.prob <- predict(model_nn, as.matrix(x_test))
pred.class <- apply(pred.prob, 1, function(x) which.max(x)-1)
# Evaluation
confMatrix <- confusionMatrix(as.factor(pred.class), as.factor(y_test))
confMatrix


```
#2.1 b) iii)
```{r}
print(pred.class)
```
I do not observe any clear pattern in the predicted values, more than many '3' predicted values.

2.1 b) iii)

The bias of the model means how far the predictions are from the target. In this case they are really far so the model has a high bias. 

2.1 b) iv)

The more epochs the more accuracy but when a point is reached, the accuracy could start decreasing, hence, we would be overfitting the model. The same happens with lower epochs, that could lead to an underfitting. So I believe adding 100 more epochs could not help in getting better results as we would probably be overfitting the neural network.

# (c) 
```{r}
model_nn <- NULL
out <- NULL
#Create, train and evaluate to choose neurons
create_model <- function(u){
  model_nn <- keras_model_sequential() %>%
    layer_dense(units=u, activation="relu", input_shape=c(7)) %>%
    layer_dense(units=5, activation="softmax")
  
  model_nn %>% compile(
    optimizer = 'adam', 
    loss = 'categorical_crossentropy',
    metrics = c('accuracy'))
 
  acc<- model_nn %>% fit(data.matrix(x_train), y_train.ohe, epochs = 100, batch_size= 32, validation_split=0.20, verbose=0)
  
 evaluation <- model_nn %>% evaluate(as.matrix(x_test), y_test.ohe)
  
  out <- list(model_nn, acc, evaluation,u)
  return(out)
}
```
#2.1 c) i)
```{r}
model_nn <- NULL
out <- NULL
out <- create_model(6)
model_nn <- out[[1]]
accur <- out[[2]]
eval <- out[[3]]
units<- out[[4]]
cat("Best model has", units, " neurons in the hidden layer \n In this model, loss:", eval[[1]], "Accuracy:", eval[[2]], "\n")

```
#2.1 c) ii)

The bias of the model is lower as the predictions are much closer to the target value.

#2.1 c) iii)
```{r}
# Plot Model Accuracy
plot(accur[["metrics"]][["accuracy"]], xlab="Epochs", ylab="Accuracy", main="Model2 Accuracy", ylim=c(0,1), col="blue", pch=20, type="o")
points(accur[["metrics"]][["val_accuracy"]], col="grey", pch=20, type="o")
legend("topright", legend=c("train", "test"),col=c("blue", "grey"), pch=c(20,20))
#Plot Model Loss
plot(accur[["metrics"]][["loss"]], xlab="Epochs", ylab="Loss", main="Model2 Loss", col="blue", pch=20, type="o")
points(accur[["metrics"]][["val_loss"]], col="grey", pch=20, type="o")
legend("topright", legend=c("train", "test"),col=c("blue", "grey"), pch=c(20,20))
```
As seen in the loss graphic, in the epoch 70 both lines converge and it does not overfit. So it could be stopped there.

# (d)
```{r}
#Confusion Matrix
# Prediction
print(model_nn)
pred.prob <- predict(model_nn, as.matrix(x_test))
pred.class <- apply(pred.prob, 1, function(x) which.max(x)-1)
# Evaluation
confMatrix <- confusionMatrix(as.factor(pred.class), as.factor(y_test))
confMatrix

cat("Best Neural Network Model", "\n")
cat("\t", "Overall accuracy", eval[[2]],"\n")

cat("\t", "Sensitivity Class1:", confMatrix$byClass[1,1], "Class2:",confMatrix$byClass[2,1] , "\n","\t","\t", "    Class3:",confMatrix$byClass[3,1], "Class4:", confMatrix$byClass[4,1], "\n" )

cat("\t", "Specificity Class1:", confMatrix$byClass[1,2], "Class2:",confMatrix$byClass[2,2] , "\n","\t","\t", "    Class3:",confMatrix$byClass[3,2], "Class4:", confMatrix$byClass[4,2], "\n" )

cat("\t", "PPV         Class1:", confMatrix$byClass[1,3], "Class2:",confMatrix$byClass[2,3] , "\n","\t","\t", "    Class3:",confMatrix$byClass[3,3], "Class4:", confMatrix$byClass[4,3], "\n" )

cat("\t", "Bal. Acc.   Class1:", confMatrix$byClass[1,11], "Class2:",confMatrix$byClass[2,11] , "\n","\t","\t", "    Class3:",confMatrix$byClass[3,11], "Class4:", confMatrix$byClass[4,11], "\n" )

```
# 2.1 d) i) 
Comparing both outputs in a) and d) it can be seen that decision tree model obtains an accurate prediction (0,97 tree model and 0,9625 neural network) with a less complex model than the neural network.

# 2.1 d) ii) 

When choosing one of these two models for deployment I would choose decision tree because it presents more or less the same accuracy but it is easier to deploy than neural networks that involve more complexity.
